The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-07 Quarto/paper.qmd
fixing quarto errors
Update Related Works

### Multi-Agent LLMs

- Among us game [@chiAMONGAGENTSEvaluatingLarge2024]
- Collective problem solving [@duLargeLanguageModels2024]
  - "analyses indicate that LLM agent groups exhibit more disagreements, complex statements, and a propensity for positive statements compared to human groups"
- Govsim [@piattiCooperateCollapseEmergence2024]
  - "In GOVSIM, a society of AI agents must collectively balance exploiting a common resource with sustaining it for future use. This environment enables the study of how ethical considerations, strategic planning, and negotiation skills impact cooperative outcomes."
- All found similar themes
  - That LLMs are capable and good at understanding the rules
  - That they can cooperate and be sneaky

### LLMs and Werewolf

- Examination of improving werewolf by LLMs [@xuLanguageAgentsReinforcement2024]
  - "our agents use an LLM to perform deductive reasoning and generate a diverse set of action candidates. Then an RL policy trained to optimize the decision-making ability chooses an action from the candidates to play in the game. Extensive experiments show that our agents overcome the intrinsic bias and outperform existing LLM-based agents in the Werewolf game."
- Werewolf Arena [@bailisWerewolfArenaCase2024]
  - Used in this paper
- Explicitly discuss how none of the exisiting LLM+Werewolf papers examine the differences/compare from a human dataset

@@

Agents participated in gameplay through a graphical user interface (GUI), depicted in Figure @fig-wwa-gui. The GUI displays the game state, including player roles, actions, and current discussion rounds, enabling monitoring of the gameplay progression.

**Figure 1.** GUI of Werewolf Arena simulation environment.

>>>>

Recent research into multi-agent large language models (LLMs) has explored their performance in various social deduction and collective problem-solving contexts. Chi et al. \[@chiAMONGAGENTSEvaluatingLarge2024] investigated LLM behavior in the popular game Among Us, revealing capabilities in understanding complex game dynamics and successfully navigating roles involving deception and cooperation. Similarly, Du et al. \[@duLargeLanguageModels2024] examined collective problem-solving scenarios, finding that LLM agent groups showed increased complexity in their interactions, more frequent disagreements, and generally positive exchanges compared to human groups. Piatti et al. \[@piattiCooperateCollapseEmergence2024], through the GovSim environment, studied how AI societies managed collective resources, demonstrating that LLM agents effectively balanced ethical considerations, strategic planning, and negotiation, further supporting the idea of their advanced cooperative and strategic capabilities.

Within the specific context of Werewolf, several studies have addressed the use of LLMs to enhance gameplay. Xu et al. \[@xuLanguageAgentsReinforcement2024] developed LLM agents that leverage deductive reasoning and reinforcement learning to optimize decision-making and gameplay strategy, outperforming existing methods. Meanwhile, Bailis et al. \[@bailisWerewolfArenaCase2024] introduced the Werewolf Arena, a framework employed in our current study. However, despite these advancements, previous research has not explicitly examined or compared LLM-driven Werewolf gameplay to authentic human interactions and strategies. Our paper specifically addresses this gap by directly comparing human gameplay, sourced from the Werewolf Among Us dataset, with simulated gameplay generated by LLMs, providing novel insights into differences and similarities in persuasion strategies and social reasoning between humans and AI.

@@

Agents participated in gameplay through a graphical user interface (GUI), depicted in @fig-wwa-gui. The GUI displays the game state, including player roles, actions, and current discussion rounds, enabling monitoring of the gameplay progression.
